Unsupervised Acquisition of Comprehensive Multiword Lexicons \\ using Competition in an $n$-gram Lattice

### 2017-01-23 tacl.tex
Comments to end of Section 3.2

@@

Work that uses collocation measures for multiword lexicon creation often assumes a small, fixed set of part-of-speech combinations \cite{Evert:Krenn:2001,Pecina10}, an assumption that breaks down entirely when considering the full range of formulaic sequences in a language. Relaxing this restriction, however, requires dealing with extreme overgeneration in terms of potential phrases; the main challenge is how to distinguish good FS in a huge collection of $n$-grams, many of which show basic statistical irregularities due to syntactic properties of the language or as the result of some relationship with a true FS. The former can be captured to some extent using a syntax-sensitive association measure like LPR, but addressing the latter requires a full, iterative model, where $n$-grams in essence compete with each other to form a compact, parsimonious formulaic vocabulary which best explains the statistics extracted from the corpus.

@@

The first step in the process is derive a set of $n$-grams and related statistics from a large, unlabeled corpus of text. Since our primary association measure is an adaption of LPR, our approach in this subsection mostly follows \newcite{Brooke15b} up until the last stage. An initial requirement of any such method is a $n$-gram frequency threshold, which for us (like Brooke et \al 2015) is 1 instance per 10 million words; we have done a manual analysis using the MWE corpus of \newcite{Schneider14a} which indicates very good (over 90\%) type-level MWE coverage using the frequency filtered $n$-gram statistics from the ICWSM blog corpus (see \secref{sec:evaluation}), excepting those highly particular proper names which we consider outside the scope of FS identification. Note that our set of $n$-grams also include gapped or non-contiguous $n$-grams; the definition of gaps is language specific---we will discuss the gaps for individual languages further in \secref{sec:evaluation}---but the general idea (and this is confirmed by empirical analysis) is that for any particular language there is a relatively small set of syntactic types of limited length that will tend to appear as a gap inside a formulaic sequence. In a part-of-speech tagged corpus, potential gaps can be identified by a part-of-speech regular expression. Putting in such a restriction is useful because it greatly lowers the number of gapped $n$-gram types, increasing precision and improving speed. Another way the number of types are lowered is by excluding punctuation and by lemmatizing.  Note that as long as the count threshold is significantly above 1, efficient extraction of all $n$-grams can be done iteratively ($n$ passes through the corpus) for both types in combination.

Once a set of relevant $n$-grams is identified (and counted), other statistics required to calculate \termdef{Lexical Predictability Ratios} (LPR) for each word in the $n$-gram are collected. LPR measures how predictable a word is in a lexical context as compared to how predictable it is given only syntactic context (over the same span of words). Formally, the LPR for some word $w_i$ in the context of a sequence of words of length $n$ which have a corresponding set of sequence of POS tags $t$, is given by:
\begin{displaymath}
LPR(w_i,w_{0,n}) = \max_{0 \leq j < k < n }{\frac{p(w_i|w_{j,k})}{p(w_i|t_{j,k})}}
\end{displaymath}

These conditional probabilities can be derived in the standard way from $n$-gram and POS $n$-gram counts. We can use the same equation for gap $n$-grams, with the caveat that quantities involving sequences which include the location where the gap occurs are derived from special gapped $n$-gram statistics, not regular $n$-gram statistics. Note that identification of the best ratio across all possible choices of context, not just the largest ,is important for longer FS, where the entire POS context alone might uniquely identify the phrase (and hence the word), resulting in a ratio of 1. It also ensures that the lower bound of LPR is 1, since the ratio for a word with no context at all is trivially 1.

In the segmentation approach of \newcite{Brooke15b}, LPR for an entire span is calculated as a product of the individual LPRs, but here we will use the minimum LPR across the words in the sequence (minLPR):
\begin{displaymath}
minLPR(w_{0,n}) = \min_{0 \leq i < n }{LPR(w_i,w_{0,n})}
\end{displaymath}

The use of minimum here should be understood in the context of competition between nodes in the lattice: minLPR for a particular $n$-gram does not reflect the \emph{overall}
degree to which an $n$-gram holds together, instead it focuses on the word which is its weakest link. For example, in the case we are evaluating the quality of \ex{be keep \gap under} as an FS, a general statistical metric might give it a high score due the strong statistical relationship between \ex{keep} and \ex{under}, but the minLPR is focused on the weaker relationship between \ex{be} and \ex{keep \gap under}.

@@

The central process of the model is to decide which nodes in an $n$-gram lattice correspond to an FS, based on the objective of maximizing the overall explanatory power of the lattice. Like probabilities, explainedness of individual nodes will vary between 0 and 1, with 1 corresponding to a node which is fully explained. It will be useful to define the explainedness of a node in terms of two functions: the rationale for having two functions is that the on/off status of a node completely changes the influence that surrounding nodes will have: when a node is off, it is generally desirable (or at least not undesirable) that connected nodes in the lattice are on, since they may serve to partially or entirely explain its existence. When a node is on, however, the presence of other turned-on nodes nearby suggests redundancy, and is not to be encouraged. Let $E_{1}$ be the explainedness of a node when it is on, corresponding to an $n$-gram which has been identified as an FS, and $E_{0}$ be the explainedness of a node when it off. The base value of $E_{0}$ is the inverse of the minLPR metric discussed above: when the minLPR metric is 1, corresponding to a case when the syntactic content predicts one of the words of the $n$-gram as well or better than the lexical context, $E_{0}$  will also be 1. The base value of $E_{1}$ is the inverse of a cost parameter $C$ and, for nodes not influenced by one of the other factors discussed below, is equivalent to the initial threshold for building a vocabulary using the minLPR association measure: all else being equal, when $minLPR(t) > C$, a node will be selected as an FS. There is no necessary upper bound on $C$, but we note here that there is a fairly small range of good choices for $C$; we have found that setting it too high ($>6$, corresponding to minimum of six times more likely than predicted by POS context alone) will restrict the model primarily to sequences with only open-class words, while setting it too low ($<3$) allows in too much random noise.

We will influence these functions by adding exponential terms, reflecting the fact we do not want to override the association measure, but rather modulate it. Let's define $E_0$, then, as the minLPR based taken to the negative power of some as yet undefined function $d_0$:
\begin{displaymath}
E_0(t) = minLPR(t)^{-d_0(t)}
\end{displaymath}
and $E_1$ is the cost parameter $C$ taken to the negative power of an analogous function $d_1$:
\begin{displaymath}
E_1(t) = C^{-d_1(t)}
\end{displaymath}
The values of $d_0$ and $d_1$ are dependent on the context of the node in the lattice, but are both 1 when a node is not being influenced by other nodes. They are also restricted to always being positive.

@@

In general, covering prefers turning on longer, covering $n$-grams since doing so explains nodes lower in the lattice. Not surprisingly, it is generally desirable to have a mechanism working in opposition, \ie one which views shorter FS as helping to explain the presence of longer ones which contain them, beyond the FS-neutral syntactic explanation provided by minLPR. \termdef{Clearing} does this by changing the $E_0$ of nodes higher in the lattice when a lower node is turned-on. The basic mechanism is similar to covering, except that counts cannot be made use of in the same way; whereas it makes sense to explain covered nodes in proportion to the counts of their covering nodes (since the counts of the covered $n$-grams are from the covering $n$-gram), in the reverse direction this logic fails. A simple but effective solution which avoids introducing parameters makes use of the minLPR values of the relevant nodes. In the most common (two node) situation, we increase $E_{0}$ of the higher node based on the ratio of the minLPR of two nodes, though only if the minLPR of the lower node is higher, which greatly restricts its effect. Generalized to consider the (rare) case of multiple clearing nodes, we define clear(t) as:
%in this paper, we consider two options: a hard clearing where any turned-on node will clear (set $C_{0}$ to 0 for) any node higher in the lattice which has a minLPR that is lower than it; and a soft 
\begin{displaymath}
clear(t) = \prod_{i=1}^{m}{ min(1,\frac{minLPR(t)}{minLPR(u_i)})}
\end{displaymath}
where $u_i$ is the $i$th clearing node of $t$. We refer to this mechanism as clearing because it tends to clear away a variety of trivial uses of common FS which many have higher LPR due to the lexical and syntactic specificity of the FS. For instance, in \figref{fig:example} if the node \ex{keep \gap under wraps}  is turned on and has a minLPR of 8, then, if the minLPR of a node such as \ex{keep \gap under wraps for} is 4, clear(t) will be 0.5. We can now define $d_{0}(t)$ as:
\begin{displaymath}
d_0(t) = cover(t) \cdot clear(t)
\end{displaymath}

>>>>

\newcommand{\LPR}{\ensuremath{\text{LPR}}}
\newcommand{\minLPR}{\ensuremath{\text{minLPR}}}

@@

Work that uses collocation measures for multiword lexicon creation often assumes a small, fixed set of part-of-speech (``POS'') combinations \cite{Evert:Krenn:2001,Pecina10}, an assumption that breaks down entirely when considering the full range of formulaic sequences in a language. Relaxing this restriction, however, requires dealing with extreme overgeneration in terms of potential phrases; the main challenge is how to distinguish good FS in a huge collection of $n$-grams, many of which show basic statistical irregularities due to syntactic properties of the language or as the result of some relationship with a true FS. The former can be captured to some extent using a syntax-sensitive association measure like LPR, but addressing the latter requires a full, iterative model, where $n$-grams in essence compete with each other to form a compact, parsimonious formulaic vocabulary which best explains the statistics extracted from the corpus.

@@

The first step in the process is to derive a set of $n$-grams and related statistics from a large, unlabeled corpus of text. Since our primary association measure is an adaption of LPR, our approach in this section mostly follows \newcite{Brooke15b} up until the last stage. An initial requirement of any such method is an $n$-gram frequency threshold, which for us (following \newcite{Brooke15b}) is 1 instance per 10 million words; we have done a manual analysis using the MWE corpus of \newcite{Schneider14a} which indicates very good (over 90\%) type-level MWE coverage using the frequency filtered $n$-gram statistics from the ICWSM blog corpus (see \secref{sec:evaluation}), excepting those highly particular proper names which we consider outside the scope of FS identification. 
% TJB: is this a subset of proper names? the "excepting ..." qualification is a bit vague
Note that our set of $n$-grams also includes gapped or non-contiguous $n$-grams. The precise definition of gaps is language specific (see \secref{sec:evaluation} for details) but the general idea is that for any particular language there is a relatively small set of syntactic types of limited length that will tend to appear as a gap inside a formulaic sequence \cite{Wasow:2002}.
% TJB: added citation and removed "this is confirmed by empirical analysis" as it begs details of the specific analysis, which I think is more of a distraction than anything else at this point
 In a POS-tagged corpus, potential gaps can be identified by a POS regular expression. This greatly lowers the number of potentially gapped $n$-gram types, increasing precision and efficiency for negligible loss of recall. We also exclude punctuation and lemmatize the corpus.  Note that as long as the count threshold is significantly above 1, efficient extraction of all $n$-grams can be done iteratively ($n$ passes through the corpus) for both types in combination. 
% TJB: this last sentence is a bit opaque

Once a set of relevant $n$-grams is identified and counted, other statistics required to calculate \termdef{Lexical Predictability Ratios} (LPR) for each word in the $n$-gram are collected. LPR measures how predictable a word is in a lexical context, as compared to how predictable it is given only syntactic context (over the same span of words). Formally, the LPR for word $w_i$ in the context of a word sequence $w_0, w_i, ..., w_{n-1}$ with POS tag sequence $t_0, t_1, ..., t_{n-1}$, is given by:
\begin{displaymath}
\LPR(w_i,w_{0,n}) = \max_{0 \leq j < k < n }{\frac{p(w_i|w_{j,k})}{p(w_i|t_{j,k})}}
\end{displaymath}
% TJB: clarify the semantics of the $w_{j,k}$ and $t_{j,k}$ notation

%These conditional probabilities can be derived in the standard way from $n$-gram and POS $n$-gram counts. 
% TJB: removed; seems to state the obvious
We use the same equation for gapped $n$-grams, with the caveat that quantities involving sequences which include the location where the gap occurs are derived from special gapped $n$-gram statistics. Note that the identification of the best ratio across all possible choices of context, not just the largest, is important for longer FS, where the entire POS context alone might uniquely identify the phrase (and hence the word), resulting in a ratio of 1. 
% TJB: is this a good thing or bad thing? slight lack of clarity here
It also ensures that the lower bound of LPR is 1, since the ratio for a word with no context is trivially 1.

In the segmentation approach of \newcite{Brooke15b}, LPR for an entire span is calculated as a product of the individual LPRs, but here we will use the minimum LPR across the words in the sequence (``\minLPR''):
\begin{displaymath}
\minLPR(w_{0,n}) = \min_{0 \leq i < n }{\LPR(w_i,w_{0,n})}
\end{displaymath}
The use of minimum here should be understood in the context of competition between nodes in the lattice: minLPR for a particular $n$-gram does not reflect the \emph{overall} degree to which an $n$-gram holds together, instead it focuses on the word which is its weakest link. For example, in the case we are evaluating the quality of \ex{be keep \gap under} as an FS, a general statistical metric might give it a high score due the strong statistical relationship between \ex{keep} and \ex{under}, but minLPR is focused on the weaker relationship between \ex{be} and \ex{keep \gap under}.

@@

The central process of the model is to decide which nodes in an $n$-gram lattice correspond to an FS, based on the objective of maximizing the overall explanatory power of the lattice. Like probabilities, explainedness of individual nodes will vary between 0 and 1, with 1 corresponding to a node which is fully explained. We define the explainedness of a node in terms of two functions, capturing the impact on surrounding nodes of a node being either off or on: when a node is off, it is generally desirable (or at least not undesirable) that connected nodes in the lattice are on, since they may serve to partially or entirely explain its existence; when a node is on, however, the presence of other turned-on nodes nearby suggests redundancy, and is not to be encouraged. Let $E_{1}$ be the explainedness of a node when it is on, corresponding to an $n$-gram which has been identified as an FS, and $E_{0}$ be the explainedness of a node when it is off. The base value of $E_{0}$ is the inverse of minLPR: when the minLPR metric is 1, corresponding to the case of the syntactic content predicting one of the words of the $n$-gram as well or better than the lexical context, $E_{0}$  will also be 1. The base value of $E_{1}$ is the inverse of a cost parameter $C$ and, for nodes not influenced by one of the other factors discussed below, is equivalent to the initial threshold for building a vocabulary using the minLPR association measure: all else being equal, when $\minLPR(t) > C$, a node will be selected as an FS. There is no necessary upper bound on $C$, but we note here that there is a fairly small range of good choices for $C$; we have found that setting it too high ($>6$, corresponding to minimum of six times more likely than predicted by POS context alone) will restrict the model primarily to sequences with only open-class words, while setting it too low ($<3$) results in too much random noise.

We weight these functions by adding exponential terms, reflecting the fact we do not want to override the association measure, but rather modulate it. We define $E_0$, then, as minLPR to the negative power of some as yet undefined non-negative function $d_0$:
\begin{displaymath}
E_0(t) = \minLPR(t)^{-d_0(t)}
\end{displaymath}
and $E_1$ is the cost parameter $C$ to the negative power of an analogous non-negative function $d_1$:
\begin{displaymath}
E_1(t) = C^{-d_1(t)}
\end{displaymath}
The values of $d_0$ and $d_1$ are dependent on the context of the node in the lattice, but are both 1 when a node is not influenced by other nodes.

@@

In general, covering prefers turning on longer, covering $n$-grams since doing so explains nodes lower in the lattice. Not surprisingly, it is generally desirable to have a mechanism working in opposition, \ie one which views shorter FS as helping to explain the presence of longer ones which contain them, beyond the FS-neutral syntactic explanation provided by minLPR. \termdef{Clearing} does this by changing the $E_0$ of nodes higher in the lattice when a lower node is turned-on. The basic mechanism is similar to covering, except that counts cannot be made use of in the same way; whereas it makes sense to explain covered nodes in proportion to the counts of their covering nodes (since the counts of the covered $n$-grams are from the covering $n$-gram), in the reverse direction this logic fails. A simple but effective solution which avoids introducing parameters makes use of the minLPR values of the relevant nodes. In the most common (two node) situation, we increase $E_{0}$ of the higher node based on the ratio of the minLPR of two nodes, though only if the minLPR of the lower node is higher, which greatly restricts its effect. Generalized to consider the (rare) case of multiple clearing nodes, we define clear(t) as:
%in this paper, we consider two options: a hard clearing where any turned-on node will clear (set $C_{0}$ to 0 for) any node higher in the lattice which has a minLPR that is lower than it; and a soft 
\begin{displaymath}
clear(t) = \prod_{i=1}^{m}{ min(1,\frac{\minLPR(t)}{\minLPR(u_i)})}
\end{displaymath}
where $u_i$ is the $i$th clearing node of $t$. We refer to this mechanism as clearing because it tends to clear away a variety of trivial uses of common FS which many have higher LPR due to the lexical and syntactic specificity of the FS. For instance, in \figref{fig:example} if the node \ex{keep \gap under wraps}  is turned on and has a minLPR of 8, then, if the minLPR of a node such as \ex{keep \gap under wraps for} is 4, clear(t) will be 0.5. We can now define $d_{0}(t)$ as:
\begin{displaymath}
d_0(t) = cover(t) \cdot clear(t)
\end{displaymath}
